---
title: "Biodiversity indices"
author: "Joel Eliason"
date: "12/28/2021"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(spatgraphs)
library(spatialOnco)
library(spatialsegregation)
```


#### Data IO/preprocessing

```{r}
DATA_PATH = '../data/nbhd_coord_schurch_2020/'
df = read_csv(paste0(DATA_PATH,'CRC_master.csv'))
pt_data = readxl::read_excel(paste0(DATA_PATH,'CRC_TMAs_patient_annotations.xlsx'))

```

```{r}
spot_eq = df %>%
  filter(spots == '15_B')
pat_eq = create_ppp(spot_eq$X,spot_eq$Y,spot_eq$ClusterName,
                    keep_types =c('Tregs','CD8+ T cells','vasculature','tumor cells'))
spot_un = df %>%
  filter(spots == '20_A')
pat_un = create_ppp(spot_un$X,spot_un$Y,spot_un$ClusterName,
                    keep_types =c('Tregs','CD8+ T cells','vasculature','tumor cells'))
```

```{r}
sg_eq = spatgraph(pat_eq,type="gabriel")
plot(sg_eq,pat_eq)
```

```{r}
sg_un = spatgraph(pat_un,type="gabriel")
plot(sg_un,pat_un)
```

* spatial Simpson index - "if we randomly pick one point of
the pattern as a focal point, and another point from the focal point’s neighbourhood,
the index describes the probability that the two are from different species."

```{r}
ntype = "delaunay"
res_eq = simpsonF(pat_eq,ntype=ntype,minusRange=F)

res_eq$S

res_un = simpsonF(pat_un,ntype=ntype,minusRange=F)

res_un$S
```

* individual species area relationship (ISAR) - This measures the number of species in the average neighbourhood of an individual of species $\tau$.

```{r}
res_eq = isarF(pat_eq,ntype=ntype,minusRange=F)

summary(res_eq)

res_un = isarF(pat_un,ntype=ntype,minusRange=F)

summary(res_un)
```

mingling index - defined as the fraction of neighbours of a different species among the neighbours of a point of species $\tau$

```{r}
res_eq = minglingF(pat_eq,ntype=ntype,minusRange=F)

summary(res_eq)

res_un = minglingF(pat_un,ntype=ntype,minusRange=F)

summary(res_un)
```
We can note that the mingling index for vasculature and tumor cells is lower in the unequal pattern than the equal pattern. This directional effect is not seen in the ISAR.
