---
title: "Spatial relationships between CD8+ T cells and Tregs"
author: "Joel Eliason"
date: "12/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(spatstat)
library(glue)
```

### Questions/Analyses

  1. Are there differences in the spatial relationships of CD8+ T cells and Tregs between patient groups? Patient stage?
  2. Is there any significant interaction of immune cells and vasculature in either patient group?
  3. At what spatial scale do we observe spatial relationships of CD8+ T cells and Tregs?
  
  
#### Data IO/preprocessing

```{r}
DATA_PATH = '../data/nbhd_coord_schurch_2020/'
df = read_csv(glue('{DATA_PATH}CRC_master.csv'))
```

```{r}
df %>%
  group_by(spots,ClusterName) %>%
  filter(ClusterName == 'Tregs' | ClusterName == 'CD8+ T cells') %>%
  summarise(n())
  
```
From this we will choose spot 15_B (almost equal ratio of Tregs and CD8+ T cells) and 20_A (more than twice as many CD8+ T cells than Tregs).

```{r}
spot_eq = df %>%
  filter(spots == '15_B')

spot_un = df %>%
  filter(spots == '20_A')
```

We can then create the spatial point processes for both spots as well:

```{r}
create_ppp = function(df,cell_types=c('Tregs','CD8+ T cells')) {
  Xmin = min(df$X) - 5
  Xmax = max(df$X) + 5
  Ymin = min(df$Y) - 5
  Ymax = max(df$Y) + 5
  
  filt = df %>%
    filter(ClusterName %in% cell_types) %>%
    mutate(ClusterName = factor(ClusterName))
  
  pat = ppp(filt$X,filt$Y,c(Xmin,Xmax),c(Ymin,Ymax))
  
  marks(pat) <- filt$ClusterName
  
  return(pat)
}

pat_eq = create_ppp(spot_eq)
pat_un = create_ppp(spot_un)

pat_un
```


### Initial visualizations

```{r}
plot(pat_eq)
plot(pat_un)
```
Quadrant counts

```{r}
plot(quadratcount(split(pat_eq)))
plot(quadratcount(split(pat_un)))

```

```{r}
plot(density(split(pat_eq)))
plot(density(split(pat_un)))
```

```{r}
miplot(pat_eq)
sapply(split(pat_eq), miplot)
```
Greater than 1 indicates clustering - so we are seeing clustering at all spatial scales in the equal pattern.

```{r}
miplot(pat_un)
sapply(split(pat_un), miplot)
```

Looks like there is still clustering at all scales, but to a lesser degree, in the unequal pattern (lower Morisita index).

```{r}
lest_eq = Lest(pat_eq, correction = "good")
plot(lest_eq)

```


```{r}
lest_un = Lest(pat_un, correction = "good")
plot(lest_un)

```


```{r}
kcross = Kcross.inhom(pat_un, 'CD8+ T cells', 'Tregs')

plot(kcross)

g = pcf(kcross)

plot(g)
```